---
title: "Assignment #5"
output:
  html_document:
    df_print: paged
---

### Set-Up

```{r}
library(readr)

AmesTrain17 <- read.csv("~/Documents/(2) Sophomore Year/Second Semester (SPR 20)/STOR 455/Assignment 3/AmesTrain17.csv")
AmesTest17 <- read.csv("/Users/alexmorton/Documents/(2) Sophomore Year/Second Semester (SPR 20)/STOR 455/CSV Files/AmesTest17.csv")
```

# PART 6

```{r}
mod1 = lm(Price~LotFrontage+LotArea+Quality+Condition+YearBuilt+YearRemodel+BasementFinSF+BasementUnFinSF+FirstSF+SecondSF+Bedroom+TotalRooms+GarageCars+GarageSF+ScreenPorchSF, data=AmesTrain17)

summary(mod1)
```

### Predicted 'Price' for Each Holdout Case

```{r}
fitPrice = predict(mod1, newdata=AmesTest17)
fitPrice
```

### Residuals for Each Holdout Case

```{r}
holdoutresid = AmesTest17$Price - fitPrice
holdoutresid
```

### Mean and Standard Deviation of Holdout Residuals

```{r}
mean(holdoutresid)
sd(holdoutresid)

mean(mod1$residuals)
sd(mod1$residuals)
```

Comparing the training and holdout residuals shows that the means vary by a decent amount (-4.499237e-16 and -3.654769), but the standard deviations remain relatively similar (28.56527 and 27.47029). So, the mean of the holdout residuals is not necessarily expected based on the training residuals.

### Poorly Predicted Holdout Cases

```{r}
hist(holdoutresid)
which(holdoutresid<=-150)
holdoutresid[c(118)]
```

Row 118 of the holdout data has a residual of -176.5973, while others have residuals contained within [-100, 100], making it especially poorly predicted by the training model.

### Cross-Validation Correlation

```{r}
crosscor = cor(AmesTest17$Price, fitPrice)

crosscor
crosscor^2
0.8713 - crosscor^2
```

The shrinkage (0.029) is subjectively OK. The training model worked very similarly for the holdout sample as the original data it was created for.

# PART 9

```{r}
AmesModFinal <- lm(log(Price)~LotArea+YearBuilt+FirstSF+SecondSF+GarageSF+ExteriorQ+ExteriorC+HeatingQC+CentralAir+BasementHt+I(FirstSF^2)+FirstSF:KitchenQ, data=AmesTrain17)
summary(AmesModFinal)

plot(AmesModFinal)

mean(resid(AmesModFinal))
```

```{r}
x = data.frame(LotFrontage=90, LotArea=11060, Quality=7, Condition=5, ExteriorQ="Gd", ExteriorC="Gd", Foundation="PConc", BasementUnFinSF=1150, BasementFBath=0, BasementHBath=0, Heating="GasA", HeatingQC="Ex", CentralAir="Y", BasementHt="Ex", KitchenQ="TA", FirstSF=1164, SecondSF=1150, Bedroom=3, FullBath=2, HalfBath=1, Fireplaces=1, GarageCars=2, GarageSF=502, GarageQ="TA", GarageC="TA", OpenPorchSF=274, TotalRooms=9, YearBuilt=1995, YearRemodel=2003, HouseStyle="2Story", BasementFin="Unf", BasementFinSF=0, ScreenPorchSF=0)

predict.lm(AmesModFinal, x, interval="prediction", level=0.95)

exp(5.291753)
exp(5.890936)
```

I chose not to make any more adjustments to the model. The mean of our residuals is already very close to 0 and there are no apparent outliers, suggesting our data fits well. We are 95% confident that some particular house with the characteristics described in 'x' costs between 198,691.40 and 361,743.70 dollars.
